---
title: "Chapter_3"
author: "Zoe"
date: "2025-09-10"
output: html_document
---

### Getting Graphic

------------------------------------------------------------------------

#### Histograms

```{r}
library(datasets)
```

```{r}
head(airquality)
```

```{r}
hist(airquality$Temp)
```

```{r}
hist(airquality$Temp,xlab="Temperature (Degrees
Fahrenheit)",main="Temperatures in New York City May 1 - September 30, 1973")
```

#### Density plots

```{r}
hist(
  airquality$Temp,
  xlab = "Temperature (Degrees Fahrenheit)",
  main = "Temperatures in New York City May 1 - September 30, 1973",
  probability = TRUE
)

lines(density(airquality$Temp), col = "blue", lwd = 2)
```

#### Bar plots

```{r}
library(MASS)
```

```{r}
head(Cars93[1:3])
```

```{r}
table(Cars93$Type)
```

```{r}
barplot(table(Cars93$Type))
```

```{r}
barplot(table(Cars93$Type),ylim=c(0,25),xlab="Type", ylab="Frequency", axis.lty="solid",space=.5)
```

```{r}
library(dplyr)
```

```{r}
USA.Cars93 <- filter(Cars93,Origin == "USA")
```

#### Grouping the bars

```{r}
females <- HairEyeColor[,,2]
```

```{r}
females
```

```{r}
color.names = c("black","grey40","grey80","white")
```

```{r}
t(females)
```

```{r}
par(mar=c(5,4,4,7)); barplot(t(females),beside=T,ylim=c(0,70),xlab="Hair Color",ylab="Frequency of Eye Color", col=color.names,axis.lty="solid"); legend("topright", rownames(t(females)), cex=0.8, fill=color.names, title="Eye Color", inset=c(-0.25,0), xpd=TRUE)
```

#### Quick Suggested Project

```{r}
males <- HairEyeColor[,,1]
```

```{r}
males
```

```{r}
color.names <- c("black", "grey40", "grey80", "white")
```

```{r}
t(males)
```

```{r}
par(mar = c(5, 4, 4, 7)); bar_midpoints <- barplot(t(males), beside = TRUE, ylim = c(0, 70), xlab = "Hair Color", ylab = "Frequency of Eye Color", col = color.names, axis.lty = "solid"); legend("topright", legend = rownames(t(males)), cex = 0.8, fill = color.names, title = "Eye Color", inset = c(-0.25, 0), xpd = TRUE)
```

#### LEGENDARY COLORS

```{r}
color.names <- c("saddlebrown",  # Brown eyes
                 "deepskyblue",  # Blue eyes
                 "goldenrod",    # Hazel eyes
                 "forestgreen")  # Green eyes
```

```{r}
brown.colors <- colors()[grep("brown", colors())]
```

```{r}
brown.colors
```

```{r}
blue.colors <- colors()[grep("blue", colors())]
```

```{r}
blue.colors
```

```{r}
hazel.colors <- colors()[grep("tan|gold|khaki|wheat|burlywood", colors())]
```

```{r}
hazel.colors
```

```{r}
green.colors <- colors()[grep("green", colors())]
```

```{r}
green.colors
```

#### QUICK SUGGESTED PROJECT: GROUPING REVISITED

```{r}
# Subset Cars93 to just Type and Origin
Type.Origin <- subset(Cars93, select = c(Type, Origin))
```

```{r}
head(Type.Origin)
```

```{r}
tbl <- table(Type.Origin$Type, Type.Origin$Origin)
```

```{r}
tbl
```

```{r}
mat <- t(tbl)
```

```{r}
cols_map <- c("USA" = "black", "non-USA" = "white")
```

```{r}
cols <- cols_map[rownames(mat)]
```

```{r}
par(mar = c(5,4,4,7)); bar_mid <- barplot(mat, beside = TRUE, col = cols, border = "black", ylim = c(0, max(mat) + 5), xlab = "Type", ylab = "Frequency", main = "Car Types by Origin"); legend("topright", legend = rownames(mat), fill = cols, border = "black", title = "Origin", inset = c(-0.25, 0), xpd = TRUE)

```

#### Pie graphs

```{r}
pie(table(Cars93$Type))
```

#### Scatterplots

```{r}
plot(airquality$Wind,airquality$Temp, pch=16,xlab = "Wind Velocity (MPH)", ylab ="Temperature (Fahrenheit)", main = "Temperature vs Wind Velocity")
```

```{r}
plot(airquality$Temp ~ airquality$Wind, pch=16,xlab = "Wind Velocity
(MPH)",ylab ="Temperature (Fahrenheit)", main = "Temperature vs Wind
Velocity")
```

#### Scatterplot matrix

```{r}
Ozone.Temp.Wind <- subset(airquality,select = c(Ozone,Temp,Wind))
```

```{r}
head(Ozone.Temp.Wind)
```

```{r}
pairs(Ozone.Temp.Wind)
```

#### Box plots

```{r}
boxplot(Temp ~ Month, data=airquality, xaxt = "n")
```

```{r}
boxplot(Temp ~ Month, data=airquality, xaxt = "n"); axis(1, at=1:5,labels=c("May","June","July","August", "September"))
```

#### Graduating to ggplot2

```{r}
library(ggplot2)
```

#### How it works

```{r}
ggplot(airquality, aes(x=Temp))
```

```{r}
ggplot(airquality, aes(x=Temp)) +
  geom_histogram()
```

##### `ggplot(data_source, aes(map data components to graph     components)) +`

##### `geom_xxx(arguments to modify the appearance of the geom) +`

##### `theme_xxx(arguments to change the overall appearance) +`

##### `labs(add axis-labels and a title)`

#### Histograms

```{r}
ggplot(airquality,aes(x=Temp)) +
  geom_histogram(binwidth = 5,color = "black",fill="grey80") + 
  theme_bw() +
  labs(x = "Temperature (Fahrenheit)",y="Frequency", title= "Temperatures in the airquality Data Frame")
```

#### Bar Plots

```{r}
ggplot(Cars93,aes(x=Type)) + 
  geom_bar()+ 
  labs(y="Frequency",title="Car Type and Frequency in Cars93")
```

#### Grouped bar plots

```{r}

```
